---
title: "lab_12"
author: "Jon Donville"
date: "04/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key
* Our usual libraries, plus the geographic ones

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)
library(lubridate)
library(ggrepel)
library(sf)
library(tidycensus)
library(tigris)

```

## Questions

**Q1.** You are interested in tracking police incidents around the West Virginia University campus over time, with the goal of surfacing interesting incidents and patterns. In particular, you want to know whether the latest incident titles are relatively new/novel compared to previous reports and whether an incident has happened before in the past 30 days.

Using [CSV data](https://raw.githubusercontent.com/dwillis/wvu-projects/master/crime-log/crime_log.csv), making the column names more readable and ensuring each datetime is separated into a date column and an hour column and also create a day_of_week column and populate that using lubridate.

Then write code that will construct a paragraph about the data described above comparing the most recent data to the previous date and the most recent previous date for the same day of the week. To do that, you will need to produce:

1. A dataframe with the crime log data loaded and the datetime column converted to the proper datatype.
2. Columns in your dataframe that contain the date, time and day of the week that each date represents (you may need to google how to do that).
3. Generate a dataframe of the most recent date's incidents.
4. Using the first title in that dataframe, find the number of incidents that occurred in the previous 30 days and write out a sentence using paste. The sentence should include the most recent date, the title of the incident and the number of times that kind of incident has occurred in the previous 30 days.

Put your sentence in the answer below, along with one other question you could ask of this data.

**A1.**
"On April 21. this crime occured: SUSPICIOUS EVENT It was the 7 time it has occurred in the last 30 days"

One other question I would ask of the data was what percentage of the crimes occurred during the operating hours of the university. 


```{r}
wv_police_data <- read_csv("https://raw.githubusercontent.com/dwillis/wvu-projects/master/crime-log/crime_log.csv")

wv_police_data_cleaned <- wv_police_data %>% 
  mutate(
    datetime=mdy_hm(datetime),
    date=date(datetime),
    time=format(datetime,
                format="%H:%M"),
    day_of_week=wday(datetime, label = TRUE, abbr = FALSE)
  ) %>% 
  arrange(desc(date)) 

latest_date <- wv_police_data_cleaned %>% slice(1)
one_month_ago <- wv_police_data_cleaned %>% 
  filter(date >= latest_date$date - 30) %>% 
  filter(title == latest_date$title) %>% 
summarise(count=n())

one_month_ago
  
  ## SENTENCE SHOULD READ. ON "LATEST DATE," "TITLE" OCCURED. IT WAS THE "COUNT, ONE MONTH AGO" TIME IT HAS HAPPENED IN THE LAST 30 DAYS


sub <- paste("On ", format(latest_date$date, format="%B %d"), ". this crime occured: ", format(latest_date$title), " It was the ", format(one_month_ago$count), " time it has occurred in the last 30 days", sep="") 
sub


```

**Q2.** Maryland produces a [CSV file](https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/Vax_ZIP_CODE_FirstSecondSingle_Totals_Raw.csv) showing the total number of vaccinations by zip code and -- interestingly -- it includes non-Maryland zip codes (as well as some zip codes that seem to be bad data). 

write a script that reads the CSV file, 

making the columns more readable and adding today's date. 

Then output the top 25 Maryland zip codes and top 25 non-Maryland zip codes based on number of people fully vaccinated, and 

write both dataframes out to CSV files, using the following name formats:

Maryland zips: maryland_zips_yyyymmdd.csv
Non-Maryland zips: non_maryland_zips_yyyymmdd.csv

Where yymmdd is from today's date. You MUST do this programmatically so that it can run every day without being updated. Be sure to add those two files to your GitHub repository.

To do that, you will need to:

1. Get rid of rows that do not have data for fully vaccinated people
2. Remove any rows where the zip code is less than five digits (there are several ways you could do this).
3. Figure out the "lowest" and "highest" Maryland zip codes.

Google the top 2-3 zip codes from both Maryland and non-Maryland results and describe them in your answer.

**A2.**
The Maryland zip codes are both in Montgomery County, and appear to be two extremely wealthy zip codes. 

The leading non-md zip code was from DC. According to median home listings, this is also a very weaolthy zip codes.

```{r}
md_vax_byzip <- read_csv("https://raw.githubusercontent.com/state-of-maryland/VaccineCSVs/master/Vax_ZIP_CODE_FirstSecondSingle_Totals_Raw.csv") %>% 
clean_names() %>% 
  mutate(date=today())

maryland_topzips <- md_vax_byzip %>% 
  filter(zip_code >= 20331, zip_code <= 21930) %>% 
  arrange(desc(fully_vaccinated)) %>% 
   head(25)

maryland_topzips

non_md_topzips <- md_vax_byzip %>% 
  filter(zip_code >= 10000, zip_code < 20331 | zip_code > 21930, zip_code <=99950) %>% 
    arrange(desc(fully_vaccinated)) %>% 
  head(25)  

non_md_topzips

write_csv(maryland_topzips, paste("maryland_zips_", today(), ".csv"))

write_csv(non_md_topzips, paste("non_md_topzips_", today(), ".csv"))


```

**Q3.** Maryland has had one of the more successful vaccination rates, but you want to understand where a significant portion of the population remains less than fully vaccinated. Make a dataframe of all Maryland zip codes from the vaccination data you used in Q2 and join it to Maryland zip code population data from the Census API. Calculate the percentage of fully vaccinated for each zip code and then make a map showing the zip codes with less than 50 percent fully vaccinated. Put a blank map with the outline of Maryland counties (which you can get from Tigris) behind it for context. Add a label to the map describing what the data is and include the date from the vaccination data (programmatically, so you can run it every day).

You WILL need to ensure that the columns you use to join the two dataframes are the same datatype and remember that join order matters.

Describe the output of the map below, and say what you think the most interesting areas are.

**A3.**
The counties that border Deleware and the ones in the northwest corner of the state are the areas where the lower vaccination rates are found. The most interesting part to me is that it seems the closer you get to urban centers like DC and Baltimore, the less you see low vaccination counties.Unsurprisingly, this trend matches with the political divides across the country, with more conservative areas found in rural locations and liberal areas near urban ones. 

```{r}
one_year <- load_variables(2019, "acs1", cache = TRUE)

```

```{r}
census_api_key("46337cfa9ca4bb84fc53936d51e61e3bdb6e6b3e")

maryland_allzips <- md_vax_byzip %>% 
  filter(zip_code >= 20331, zip_code <= 21930) %>%
   mutate(zip_code = as.character(zip_code)) %>% 
  arrange(desc(fully_vaccinated)) 

md_pop <- get_acs(geography = "zcta", variables = "B01001_001", state="MD", year = 2019, geometry = TRUE) %>% 
  rename(zip_code = GEOID) %>% 
  mutate(zip_code = as.character(zip_code))

current <- maryland_allzips %>% summarize(max(date))
marylandcurrent <- maryland_allzips %>% filter(date == current[[1]])

md_vax_bypop <- md_pop %>% 
  left_join(maryland_allzips) %>% 
  mutate(percent_vaxxed = (fully_vaccinated/estimate)*100) %>% 
filter(percent_vaxxed >= 0 , percent_vaxxed <= 50) %>% 
  arrange(desc(percent_vaxxed))

md_counties <- counties() %>% filter(STATEFP == "24")

get_date <- maryland_allzips %>% 
  slice(1)

ggplot() +
  geom_sf(data=md_counties)+
   geom_sf(data=md_vax_bypop, aes(fill=percent_vaxxed))+
scale_fill_viridis_c(option = "plasma", trans = "sqrt") +
  theme_void() +
  labs(title = paste("COVID-19 vaccination rates as of ", current[[1]], sep=""))
```


```{r}

```
