---
title: "Reverse Engineering Project"
author: "Mythili Devarakonda, Varun Shankar, Jon Donville"
date: "3/5/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban (WRAL, North Carolina)](https://www.wral.com/in-session-campaign-donations/17716104/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(refinr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 


This data comes from campaign finance data. The main sections we focused on were name-clean (Name of donater), committee-name (who the money was for), amount (how much the donation was for). We had to do some mild cleaning of the data with the names and then changed the format of the date to be better served.

```{r}
# Loaded required data by reading csv file
nc_finance_data <- read_csv("data/donors0103-0701.csv")
View(nc_finance_data)
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Add in donations to the two main political parties and to a handful of other large fundraising vehicles for state legislative races, and the in-session giving topped $3.4 million.”

* **Analysis summary**:I modified the date column's type. I filtered using that new date format and calculated the total amount, and then summarized it to get the right result of $3.4 million.

```{r}
# Put code to reverse engineer sentence here
# Cleaning the date column and modifying the amount column's datatype from character to number
nc_finance_data <- nc_finance_data %>% 
  mutate(date = mdy(date))  
nc_finance_data <- nc_finance_data %>% mutate(amount = as.numeric(parse_number(amount)))
# Filtering the dates between May and June 2018 
major_donations<-nc_finance_data %>% filter( date >= '2018-05-16' & date <= '2018-06-29' )
# Summarizing the total amount in the above date range
major_donations %>% summarise(
  total_amount = sum(amount)
)
  
```

### Sentence 2

* **Sentence text**: “Berger, R-Rockingham, brought in about $60,000 that week. House Speaker Tim Moore's campaign account collected more than $72,000. Both men declined interview requests on these issues.”

* **Analysis summary**:After using the changed date format, I filtered between 5/9 and 5/16 (the week before the session started) and filtered for Berger and Moore's committee names to find the correct amounts they were given in that time frame.


```{r}
# Put code to reverse engineer sentence here
# Filtering Berger donations by date and committee_name
berger_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-09' & date <= '2018-05-16' & committee_name == "PHILIP E BERGER COMM") 
# Summarizing the total loan amount of Berger donations
berger_donations %>% 
summarize(total_loans_amount = sum(amount))
#Filtering Moore donations by date and committee_name
moore_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-09' & date <= '2018-05-16' & committee_name == "FRIENDS OF TIM MOORE") 
#Summarizing the total loan amount of Moore donations
moore_donations %>% 
summarize(total_loans_amount = sum(amount))
  # Display results of code below this codeblock
```



### Sentence 3

* **Sentence text**: “R.H. Barringer's board chair, Jasie Barringer, donated another $8,200 in-session, and Teresa Craig kicked in another $5,200 to Berger. The North Carolina Beer & Wine Wholesalers Association PAC gave $18,400 during the week before session and on opening day.”

* **Analysis summary**: I filtered by date and by name for each donation. I did each one in their own data frame to keep it as clear as possible. My results matched the article's findings.

```{r}
# Put code to reverse engineer sentence here
#Read cleaned csv file
nc_finance_data_cleaned <- read.csv("data/nc_finance_data_cleaned.csv") 
# Clean it to change the datatype of the amount column
nc_finance_data_cleaned <- nc_finance_data_cleaned %>%  mutate(amount = as.numeric(parse_number(amount)))
# Filtering for Barringer loans by name
barringer_donations<-nc_finance_data_cleaned %>% 
  filter(str_detect(name_clean, "BARRINGER")) 
# Filtering for Barringer loans by date
barringer_donations_newdates <- barringer_donations %>% mutate(date = mdy(date)) %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29')
# Filtering for Craig loans by name
craig_donations <- nc_finance_data_cleaned %>%  
  filter(str_detect(name_clean, "TERESA S CRAIG"))
# Filtering for Craig loans by date
craig_donations_newdates <- craig_donations %>% mutate(date = mdy(date)) %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29')
#Filtering for beer and wine donations by name
beer_andwine_donations <- nc_finance_data_cleaned %>% 
  filter(str_detect(name_clean, "BEER & WINE"))
#Filter for beer and wine donations by date
beer_andwine_newdates <- beer_andwine_donations %>% mutate(date = mdy(date)) %>% 
  filter(date >= '2018-05-09' & date <= '2018-05-16') 
# Summarizing each donation to find total amount
beer_andwine_newdates %>% 
  summarise(
  total_amount = sum(amount))
craig_donations_newdates %>% 
  summarise(
  total_amount = sum(amount))
barringer_donations_newdates %>% 
  summarise(
  total_amount = sum(amount))
```

### Sentence 4

* **Sentence text** : “Senate President Pro Tem Phil Berger's campaign alone raised nearly $222,000 during session. Twenty-three people gave him the maximum allowed: $5,200.”

* **Analysis summary**: By filtering by the date and committee name and then summarizing the total amount of donations, I was able to confirm Berger got just under 222k in donations during the session. 
I then did the same but also added another filter requirement, that the amount donated was the maximum allowed of 5200, and I got 23 results, confirming what the story said.

```{r}
# Put code to reverse engineer sentence here
#Filtering by name and committee_name for Berger donations
berger_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-16' & date <= '2018-06-29' & committee_name == "PHILIP E BERGER COMM") 
#Summarizing the total loan amount
berger_donations %>% 
summarize(total_loans_amount = sum(amount))
#Filtering for the maximum allowed donations
berger_max_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-16' & date <= '2018-06-29' & committee_name == "PHILIP E BERGER COMM" & amount == 5200) 
#Displaying the maximum allowed Berger donations
berger_max_donations
```


### Sentence 5

* **Sentence text**: "The co-chief executives at a Georgia firm that brokers tax credits put $41,000 into the campaign accounts of North Carolina General Assembly members during the last regular legislative session."

* **Analysis summary**: My analysis showed that during the session from May 16 to June 29, the two executives gave a combined $41,250. I am not sure if the authors of the article rounded their results for readability, but I had the total about 250 dollars different than the article. It also seems that the article made a mistake in saying that they donated these figures during the last REGULAR legislative session. According to ballotpedia (https://ballotpedia.org/2018_North_Carolina_legislative_session), the last regular session at the time of the article went from January 10 - July 4. The two executives donated significantly more than 41k during that time.

```{r}
# Put code to reverse engineer sentence here
# Filter for Delmer donations, by date, name, and make sure the donations were given to candidates
delmer_donations<-nc_finance_data_cleaned %>% 
  mutate(date = mdy(date)) %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29') %>% 
  filter(str_detect(cleaned_name, "Delmer")) %>% 
  filter(str_detect(source, "candidates")) %>% 
  filter(str_detect(transaction_type, "Individual"))
# Filter for Strobel donations, by date, name, and make sure the donations were given to candidates
strobel_donations<-nc_finance_data_cleaned %>% 
  mutate(date = mdy(date)) %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29') %>% 
  filter(str_detect(cleaned_name, "Strobel")) %>% 
  filter(str_detect(source, "candidates")) %>% 
  filter(str_detect(transaction_type, "Individual"))
  
# Summarize and display donations for Delmer and Strobel
delmer_total <- delmer_donations %>% 
  summarise(
  total_amount = sum(amount))
strobel_total <- strobel_donations %>% 
  summarise(
  total_amount = sum(amount))
# Adding Strobel and Delmer's donations for a combined value
combined_total <- delmer_total + strobel_total
```

-30-
