---
title: "lab_06"
author: "Jonathan Donville"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* [install Tabula](https://tabula.technology/)
* run existing code as directed (look for **Task**).
* write code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes in the lab_06 folder and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Get Our PDF

We'll be working with the [Maryland Statistical Handbook](https://planning.maryland.gov/MSDC/Documents/md-statistical-handbook.pdf) from the state Department of Planning. It's not tiny (44MB), and you'll want to download it to a place you'll remember (like your Downloads folder).

**Task** Download the file from the above link and save it to your Downloads folder.

## Setup Tabula

**Task** Start Tabula and load the PDF, then extract a table

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the Statistical Handbook file in your Downloads folder and click "open", and then click the "Import button" in Tabula. This will take a minute or two.

Let's go to page 30 of the PDF, Table 2A, "International Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". This is _net migration_, meaning it shows the total number of people moving to Maryland and its regions/jurisdictions _minus_ the total number of people leaving the same, divided into 12-month periods from July to July. In Tabula, draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV (it should be called `tabula-md-statistical-handbook.csv` by default) to your lab_06/data folder.

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)):

**Task** Write code to read the CSV file into a dataframe called `international_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores.

```{r}
# Write the code to load the CSV file here
international_migration <- read_csv("tabula-md-statistical-handbook.csv", skip=2, col_names=c("State.Jurisdiction", "July_2011", "July_2012", "July_2013", "July_2014", "July_2015", "July_2016", "July_2017", "July_2018", "July_2019", "total"))

```

Add a column for the type of migration ("international") and populate it:

**Task** Write the code to add a `migration_type` column to the data and give it the value of "international"

```{r}
# Write the code to add the column here
international_migration_withtype <- international_migration %>% 
  mutate(migration_type =("International"))


```

## Answer questions

**Q1** Which region and county/city _below_ the state level accounted for the largest percentage of international migration for the entire time period? You'll need to add and populate a column representing percent of total using `mutate`.
**A1** [The Suburban Washington region was the highest percentage for region. Montgomery county was highest for city/county. ]

```{r}
# Write the code to produce your answer here
international_migration_withtype_andpercent <- international_migration_withtype %>% 
  mutate(percent_of_total = ((total/198996)*100))
```

**Q2** Write a sentence or two that describes the data you produced in A1. Try to convey the most important idea.
**A2** [The most important idea to me is that there is considerably more internation migration from Washington than there is from Baltimore. Another takeaway is that the migration is very concentrated between the Suburban Washington and Baltimore areas, with very small percentages coming from cities and counties that are not included in those regions.]

**Q3** Which region & jurisdiction had the largest percentage change (either negative or positive) for international migration between July 2018 and July 2017? The formula for percentage change is easy to remember: (New-Old)/Old.
**A3** [The biggest change was -76.9% in Calvert and the biggest regional change was in Southern Maryland.]

```{r}
# Write the code to produce your answer here
international_migration_withtype_andjuly <- international_migration_withtype_andpercent %>% 
mutate(percent_change = (((July_2018-July_2017)/July_2017)*100)) %>% 

group_by(State.Jurisdiction) %>% 
arrange()
```

**Q4** What's your best guess as to why these changes occurred, and in those places in particular?
**A4** [I could not find any significant indicators of why this change might have happened. My best guess is that both areas had relatively small sample sizes, so changes in count can be overemphasized by percentage change.]


## Back to Tabula

**Task** Extract second table from Tabula

Let's go to page 31 of the PDF, Table 2B, "Domestic Migration for Maryland's Jurisdictions, July 1, 2010 to July 1, 2019". In Tabula, hit the "Clear All Selections" button and then draw a box around that table's border and click the "Preview & Export Extracted Data" button. It should look pretty clean. Let's export that CSV to your lab_06/data folder. (and rename it to `tabula-md-statistical-handbook-domestic.csv`).

## Cleaning up the data in R

Let's load it into R, and in doing so we'll skip the first two rows and add our own headers that are cleaner:

**Task** Write code to read the CSV file into a dataframe called `domestic_migration`. As in the pre_lab, you may have to skip 1 or more rows and you may want to supply easier to use column names that use underscores. Your work on the international file should serve as a guide.

```{r}
# Write the code to load the CSV file here
domestic_migration <- read_csv("data/tabula-md-statistical-handbook-domestic .csv", skip=2, col_names=c("State.Jurisdiction", "July_2011", "July_2012", "July_2013", "July_2014", "July_2015", "July_2016", "July_2017", "July_2018", "July_2019", "total"))

```

Add a column for the type of migration ("domestic") and populate it:

**Task** Write the code to add a `migration_type` column to the data and populate it with the value of "domestic"

```{r}
# Write the code to add the column here
domestic_migration_withtype <- domestic_migration %>% 
  mutate(migration_type =("Domestic"))

```

## Answer questions
**Q5** Which Maryland individual jurisdiction saw the largest net decrease in domestic migration overall?
**A5** [The Suburban washington region had the largest decrease, but I am not sure if that counts as an individual jurisdiction. If not, Baltimore City would have the largest decrease.]

```{r}
# Write the code to produce your answer here
domestic_migration_withtype %>% 
  group_by(State.Jurisdiction) %>% 
  summarise(total) %>% 
  arrange((total))

```

**Q6** How many regions & jurisdictions had net positive migration for July 2017, July 2018 and July 2019 (not just individually but all three together)?
**A6** [14 regions and jurisdictions have this migration.]

```{r}
# Write the code to produce your answer here
domestic_migration_withtype %>% 
  filter(July_2017 > 0 &
           July_2018 > 0 &
           July_2019 > 0) 




```

**Q7** How would you describe this data? Is there a county or region that stands out, and why?
**A7** [The data is a little messy in the sense that many of these jurisdictions have negative numbers in other years except the ones we cherrypicked. One that stands out is Charles, which never reports negative numbers across this data set.]
