---
title: "lab_01"
author: "Jonathan Donville "
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data.

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** Rows: 33,336, columns: 58

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

Use the pre-lab and your reference notebook as examples.

**Q2** Which West Virginia county got the most loans? How many?
**A2** [Kanawha County got the most loans]

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** [Wirt county is the actual West Virginia county with the fewest loans. They received 36 loans]

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange((count_loans))

```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** [The dataset has Palm Beach and Allegheny as counties. These are not actual counties in West Virginia and therefore that explains how they have so few recorded loans. For the actual counties, Wirt county is the smallest county in West Virginia by population, while Kanawha county has the highest largest population. ]


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** [The largest loan was given to the Greenbrier Hotel Corporation]

```{r}
wv_ppp_loans %>%
  arrange(desc(amount))

```

**Q6** What West Virginia politician (with his family) owns the company identified in question 5? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** [West Vriginia Governor Jim Justice and his family own the hotel corporation. The source is this article from propublica: https://www.propublica.org/article/west-virginia-greenbrier-governor-jim-justice-little-trump]

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** [The most relevent finding was that the average loan was actually quite small. While a few very large loans pushed the mean loan amount higher, the median loan amoung of just over 20,000 dollars shows that most businesses were getting relatively little help from the government. As with the Maryland data, it is also notable just how small some of the loans are, with a minimum loan of 165 dollars, some businesses were essentially getting no help from the government.]
```{r}
  wv_ppp_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```
